Your Paper Title

### 2025-03-31 paper/paper.qmd
wrote part of data section

---
title: "Your Paper Title"
subtitle: "Understanding Temperature's Role in Earlier Cherry Blossom Blooming Times in Kyoto, Japan"
author:
    - Luca Carnegie
    - Chris Lu
    - Russell Luchin
    - Randall Ni
abstract:
    This is the abstract of your paper. Write a brief summary of your research,
    including your objectives, methods, key results, and main conclusions.
    Keep it concise but informative.
format:
    pdf:
        documentclass: article
        papersize: letter
        toc: true
        number-sections: true
        colorlinks: true
        fig-cap-location: bottom
        tbl-cap-location: top
        fontsize: "12"
        geometry: "margin=1.2in" # print margins (change if wanting to)
bibliography: references.bib
execute:
    echo: false  # Hide code by default
    warning: false
    message: false
    jupyter: "python"
nocite: "@*" # display all references by default
thanks: "Code and workflow available at: https://github.com/lcarnegie/sakura" 
---

@@

In this paper, we analyse and model the nearly 1200-year record compiled by @Aono2010 along modern data compiled by the Japanese Meteorological Agency (JMA) [@JMA_sakura] [@JMA_temp], collected together by @Cookson. In particular, we seek to answer the question of whether warmer temperatures lead to earlier blooming times. @sec-lit-review does a broad review of existing work surrounding cherry blossoms, their significance, and efforts to understand their behaviour related to climate. @sec-data covers the data sources, methodology, and conducts an exploratory analysis of the dataset. @sec-model proposes a model of cherry blossom flowering day as a function of mean March temperature in Kyoto and lists its results. @sec-discussion critically examines the data and results, comes to several conclusions about the dataset, and proposes new areas of exploration. This analysis highlights the need for... [FILL IN AT END]

@@

@@

@@

>>>>

---
title: "Your Paper Title"
subtitle: "Understanding Temperature's Role in Earlier Cherry Blossom Blooming Times in Kyoto, Japan"
author:
    - Luca Carnegie
    - Chris Lu
    - Russell Luchin
    - Randall Ni
abstract:
    This is the abstract of your paper. Write a brief summary of your research,
    including your objectives, methods, key results, and main conclusions.
    Keep it concise but informative.
format:
    pdf:
        documentclass: article
        papersize: letter
        toc: true
        number-sections: true
        colorlinks: true
        fig-cap-location: bottom
        tbl-cap-location: top
        fontsize: "12"
        geometry: "margin=1.2in" # print margins (change if wanting to)
bibliography: "references.bib"
execute:
    echo: false  # Hide code by default
    warning: false
    message: false
    jupyter: "python"
nocite: "@*" # display all references by default
thanks: "Code and workflow available at: https://github.com/lcarnegie/sakura" 
---

```{python}
# Import necessary libraries 
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
import matplotlib.dates as mdates
import seaborn as sns
import statsmodels.api as sm
import numpy as np
from matplotlib.colors import LinearSegmentedColormap
from matplotlib.ticker import MaxNLocator
from datetime import datetime

# Load the historical data
historic_bloom = pd.read_csv("../data/analysis_data/historic_bloom.csv")

# Load the modern data
modern_bloom = pd.read_csv("../data/analysis_data/modern_bloom.csv")

# Truncate temperature to 1 decimal point
historic_bloom['avg_temp_march_c'] = historic_bloom['avg_temp_march_c'].round(1)
modern_bloom['avg_temp_march_c'] = modern_bloom['avg_temp_march_c'].round(1)

```

@@

We seek to answer the question of whether warmer temperatures lead to earlier blooming times. To do this, we analyse and model the nearly 1200-year record compiled by @Aono2010 in tandem with modern data compiled by the Japanese Meteorological Agency (JMA), collected and compiled together by @Cookson.@sec-lit-review does a broad review of existing work surrounding cherry blossoms, their significance, and efforts to understand their behaviour related to climate. @sec-data covers the data sources, methodology, and conducts an exploratory analysis of the dataset. @sec-model proposes a model of cherry blossom flowering day as a function of mean March temperature in Kyoto and lists its results. @sec-discussion critically examines the data and results, comes to several conclusions about the dataset, and proposes new areas of exploration. This analysis highlights the need for... \[FILL IN AT END\]

@@

To investigate the relationship between mean March temperature and blooming date in Kyoto, historical data on cherry blossom blooming dates [@Aono2010], as well as data collected through modern measurement techniques [@JMA_sakura] [@JMA_temp] were analyzed. @Cookson provided both datasets assembled in a single repository, though the modern data sources were last updated in 2020.

All of the data was transformed in Python [@python], using the `pandas` [@pandas] and `numpy` [@numpy] packages. For this analysis, the average (mean) march temperature, bloom date, day of year of blooming and year were made of interest.

In the historical data, @Aono2010 reconstructs the mean march temperature in Kyoto for some of the years, based on the blooming date, but also provided observed temperatures taken from local newspapers for a given year when possible. We prioritized used observed temperatures, but if none were available, then the reconstructed one was used. This created a dataset of 827 observations, from the year 812-2015. A sample of the table can be seen in @tbl-historical.

With the modern data, @Cookson combined average monthly temperatures and blooming date in a single dataset for several cities across Japan. We filtered this data to solely Kyoto, with average march temperatures to match the historical data. This created a dataset with 67 observations from the year 1953-2015. A sample can be seen in @tbl-modern.

In our data exploration, the `matplotlib` [@matplotlib] and `seaborn` packages were used to visualize the data.

```{python}
#| label: tbl-historical
#| tbl-cap: "Historical Cherry Blossom Blooming"

# Rename columns for better readability
historic_bloom_renamed = historic_bloom.rename(columns={
    'year': 'Year',
    'flower_date': 'Bloom Date',
    'flower_doy': 'Day of Year',
    'avg_temp_march_c': 'March Temp (°C)'
})
# Display the table without .style.set_caption()
display(historic_bloom_renamed.head())
```

```{python}
#| label: tbl-modern
#| tbl-cap: "Modern Cherry Blossom Blooming"

# Rename columns for better readability
modern_bloom_renamed = modern_bloom.rename(columns={
    'year': 'Year',
    'flower_date': 'Bloom Date',
    'flower_doy': 'Day of Year',
    'avg_temp_march_c': 'March Temp (°C)'
})

# Display the table without .style.set_caption()
display(modern_bloom_renamed.head())
```

The datasets were first explored chronologically, beginning with the historical data and ending with the modern data.

@@

asl;dkfj;laksdfjl;asjdf;lasjdfljasl;fj;lasjdf

@@

alsdhj;lasjfljas;lfja;lsjf;lasjdf;
